### 2019-05-14 AnalysisCode/ms1_analyses.Rmd
limno data import functions and code

---
title: 'Q1: Are lake and terrestrial primary productivity coherent?'
author: "Jonathan Walter, Grace Wilkinson, Rachel Fleck, Michael Pace"
date: "4/17/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(aqts)
library(wsyn)
library(rgdal)
library(LAGOSNE)

rm(list=ls())

```

This document organizes for openness and reproducibility analyses of the temporal coherence of interannual variation in lake primary productivity with terrestrial primary productivity in the landscape surrounding the lake.

# Data important and formatting

```{r limno import and formatting, echo=TRUE, cache=TRUE}

lakes_to_get<-read.csv("~/GitHub/AquaTerrSynch/AnalysisCode/lakes20yrs_chla.csv")$x

lagoslakes.raw<-makeLAGOSannualts(lakes_to_get, tsvars=c("chla"), minmos=3, minobs=3)
lagoslakes.cln<-cleanAnnualts(lagoslakes.raw)

rm(lakes_to_get)

MNMPCAlakes.raw<-makeMNMPCAannualts(dir="/Users/jonathanwalter/Box Sync/NSF EAGER Synchrony/Data/LAGOS Extended/MN_MPCA",
                                    minmos=3, minobs=3)
MNMPCAlakes.raw<-fixNamestoLAGOS(MNMPCAlakes.raw,"Minnesota MPCA")
MNMPCAlakes.cln<-cleanAnnualts(MNMPCAlakes.raw)

WIDNRlakes.raw<-makeWIDNRannualts(dir="/Users/jonathanwalter/Box Sync/NSF EAGER Synchrony/Data/LAGOS Extended/WI_DNR",
                                   minmos=3, minobs=3)
WIDNRlakes.raw<-fixNamestoLAGOS(WIDNRlakes.raw, "Wisconsin DNR")
WIDNRlakes.cln<-cleanAnnualts(WIDNRlakes.raw)

analysislakes<-list(lakeinfo=rbind(lagoslakes.cln$lakeinfo,MNMPCAlakes.cln$lakeinfo,WIDNRlakes.cln$lakeinfo),
                    lakedata<-c(lagoslakes.cln$lakedata, MNMPCAlakes.cln$lakedata, WIDNRlakes.cln$lakedata))

any(duplicated(analysislakes$lakeinfo$lagoslakeid)) #no duplicates

```

```{r AVHRR data import, echo=TRUE}

#Add the AVHRR data

```