### 2019-05-20 AnalysisCode/ms1_analyses.Rmd
coherence analyses

---
title: 'Q1: Are lake and terrestrial primary productivity coherent?'
author: "Jonathan Walter, Grace Wilkinson, Rachel Fleck, Michael Pace"
date: "4/17/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(aqts)
library(wsyn)
library(rgdal)
library(LAGOSNE)
library(raster)
library(MuMIn)
library(dplyr)
library(nlme)

rm(list=ls())

```

This document organizes for openness and reproducibility analyses of the temporal coherence of interannual variation in lake primary productivity with terrestrial primary productivity in the landscape surrounding the lake.

# Data important and formatting

```{r limno import and formatting, echo=TRUE, cache=TRUE}

lakes_to_get<-read.csv("~/GitHub/AquaTerrSynch/AnalysisCode/lakes20yrs_chla.csv")$x

lagoslakes.raw<-makeLAGOSannualts(lakes_to_get, infovars=c("lake_area_ha", "lake_perim_meters", "nhd_ftype", "hu4_zoneid"), tsvars=c("chla"), 
                                  minmos=3, minobs=3, timespan=c(1989,2018))
lagoslakes.cln<-cleanAnnualts(lagoslakes.raw)

rm(lakes_to_get)

MNMPCAlakes.raw<-makeMNMPCAannualts(dir="/Users/jonathanwalter/Box Sync/NSF EAGER Synchrony/Data/LAGOS Extended/MN_MPCA",
                                    infovars=c("lake_area_ha", "lake_perim_meters", "nhd_ftype", "hu4_zoneid"), minmos=3, minobs=3, timespan=c(1989,2018))
MNMPCAlakes.raw<-fixNamestoLAGOS(MNMPCAlakes.raw,"Minnesota MPCA")
MNMPCAlakes.cln<-cleanAnnualts(MNMPCAlakes.raw)

WIDNRlakes.raw<-makeWIDNRannualts(dir="/Users/jonathanwalter/Box Sync/NSF EAGER Synchrony/Data/LAGOS Extended/WI_DNR",
                                  infovars=c("lake_area_ha", "lake_perim_meters", "nhd_ftype", "hu4_zoneid"), minmos=3, minobs=3, timespan=c(1989,2018))
WIDNRlakes.raw<-fixNamestoLAGOS(WIDNRlakes.raw, "Wisconsin DNR")
WIDNRlakes.cln<-cleanAnnualts(WIDNRlakes.raw)

analysislakes<-list(lakeinfo=rbind(lagoslakes.cln$lakeinfo,MNMPCAlakes.cln$lakeinfo,WIDNRlakes.cln$lakeinfo),
                    lakedata=c(lagoslakes.cln$lakedata, MNMPCAlakes.cln$lakedata, WIDNRlakes.cln$lakedata))

any(duplicated(analysislakes$lakeinfo$lagoslakeid)) #no duplicates

```

```{r AVHRR data import, echo=FALSE, cache=TRUE}

dbuff<-calcBufferDist(sa=analysislakes$lakeinfo$lake_area_ha, ltype=analysislakes$lakeinfo$nhd_ftype, minbuff=2500)

accndvi<-stack("/Users/jonathanwalter/Box Sync/NSF EAGER Synchrony/Data/SatelliteData/accvi1989_2018.img")
maxndvi<-stack("/Users/jonathanwalter/Box Sync/NSF EAGER Synchrony/Data/SatelliteData/mxvi1989_2018.img")

nlcdwater<-raster("/Users/jonathanwalter/Box Sync/NSF EAGER Synchrony/Data/SatelliteData/landcover/nlcd2011_cls11.img")
watermask<-nlcdwater<0.05

accndvi<-mask(accndvi,watermask,maskvalue=0) #mask out pixels that have too much water in them
maxndvi<-mask(maxndvi,watermask,maskvalue=0)

analysislakes<-addAVHRRannualts(analysislakes, accndvi, dbuff)
analysislakes<-addAVHRRannualts(analysislakes, maxndvi, dbuff)

save.image("/Users/jonathanwalter/Box Sync/NSF EAGER Synchrony/Data/RData files/ms1_analysis_inprogress1.RData")

```

```{r examine missing values, echo=TRUE, cache=FALSE}
any(sapply(analysislakes$lakedata, function(x){any(is.infinite(x))}))
any(sapply(analysislakes$lakedata, function(x){any(is.na(x))}))
which(sapply(analysislakes$lakedata, function(x){any(is.na(x))}))
analysislakes$lakeinfo[which(sapply(analysislakes$lakedata, function(x){any(is.na(x))})),]

image(accndvi)
points(lakepts.prj[which(sapply(analysislakes$lakedata, function(x){any(is.na(x))})),])

dbuff[which(sapply(analysislakes$lakedata, function(x){any(is.na(x))}))]

analysislakes$lakeinfo<-analysislakes$lakeinfo[!sapply(analysislakes$lakedata, function(x){any(is.na(x))}),]
analysislakes$lakedata<-analysislakes$lakedata[!sapply(analysislakes$lakedata, function(x){any(is.na(x))})]
```

```{r compute coherences, echo=TRUE, cache=FALSE}

source("~/GitHub/AquaTerrSynch/AnalysisCode/bandtest_coh.R")

tsranges<-rbind(c(2,4),c(4,Inf),c(2,Inf))

coh.chlaXaccndvi<-NULL
coh.chlaXmaxndvi<-NULL

for(lind in 1:length(analysislakes$lakedata)){
  lakedat.ii<-cleandat(analysislakes$lakedata[[lind]], as.numeric(colnames(analysislakes$lakedata[[lind]])), clev=5)$cdat
  chlaXaccndvi<-coh(lakedat.ii[1,], lakedat.ii[2,], as.numeric(colnames(analysislakes$lakedata[[lind]])),
                    norm="powall", sigmethod="fast", nrand=10000)
  chlaXmaxndvi<-coh(lakedat.ii[1,], lakedat.ii[3,], as.numeric(colnames(analysislakes$lakedata[[lind]])),
                    norm="powall", sigmethod="fast", nrand=10000)
  for(rind in 1:nrow(tsranges)){
    chlaXaccndvi<-bandtest.coh(chlaXaccndvi, tsranges[rind,])
    chlaXmaxndvi<-bandtest.coh(chlaXmaxndvi, tsranges[rind,])
  }
  coh.chlaXaccndvi<-rbind(coh.chlaXaccndvi, c(t(as.matrix(chlaXaccndvi$bandp[,3:5]))))
  coh.chlaXmaxndvi<-rbind(coh.chlaXmaxndvi, c(t(as.matrix(chlaXmaxndvi$bandp[,3:5]))))

}

coh.chlaXaccndvi<-as.data.frame(coh.chlaXaccndvi)
coh.chlaXmaxndvi<-as.data.frame(coh.chlaXmaxndvi)

colnames(coh.chlaXaccndvi)<-paste0("accndvi",c("p.ts1","phi.ts1","coh.ts1","p.ts2","phi.ts2","coh.ts2","p.ts3","phi.ts3","coh.ts3"))
colnames(coh.chlaXmaxndvi)<-paste0("maxndvi",c("p.ts1","phi.ts1","coh.ts1","p.ts2","phi.ts2","coh.ts2","p.ts3","phi.ts3","coh.ts3"))

coh.chlaXaccndvi$lagoslakeid<-analysislakes$lakeinfo$lagoslakeid
coh.chlaXmaxndvi$lagoslakeid<-analysislakes$lakeinfo$lagoslakeid

```

```{r summarize coherences, echo=TRUE, cache=FALSE}
#short timescales
hist(coh.chlaXaccndvi$accndvicoh.ts1)
hist(coh.chlaXmaxndvi$maxndvicoh.ts1)

quantile(coh.chlaXaccndvi$accndvip.ts1)
quantile(coh.chlaXmaxndvi$maxndvip.ts1)

alpha=0.05
sum(coh.chlaXaccndvi$accndvip.ts1<alpha)/nrow(coh.chlaXaccndvi)
sum(coh.chlaXmaxndvi$maxndvip.ts1<alpha)/nrow(coh.chlaXmaxndvi)

print(coh.chlaXaccndvi$accndviphi.ts1[coh.chlaXaccndvi$accndvip.ts1<alpha]/pi) #only pattern is that lakes don't lead the landscape
print(coh.chlaXmaxndvi$maxndviphi.ts1[coh.chlaXmaxndvi$maxndvip.ts1<alpha]/pi)

#long timescales
hist(coh.chlaXaccndvi$accndvicoh.ts2)
hist(coh.chlaXmaxndvi$maxndvicoh.ts2)

quantile(coh.chlaXaccndvi$accndvip.ts2)
quantile(coh.chlaXmaxndvi$maxndvip.ts2)

alpha=0.05
sum(coh.chlaXaccndvi$accndvip.ts2<alpha)/nrow(coh.chlaXaccndvi)
sum(coh.chlaXmaxndvi$maxndvip.ts2<alpha)/nrow(coh.chlaXmaxndvi)

print(coh.chlaXaccndvi$accndviphi.ts2[coh.chlaXaccndvi$accndvip.ts2<alpha]/pi)
print(coh.chlaXmaxndvi$maxndviphi.ts2[coh.chlaXmaxndvi$maxndvip.ts2<alpha]/pi)

#all timescales
hist(coh.chlaXaccndvi$accndvicoh.ts3)
hist(coh.chlaXmaxndvi$maxndvicoh.ts3)

quantile(coh.chlaXaccndvi$accndvip.ts3)
quantile(coh.chlaXmaxndvi$maxndvip.ts3)

alpha=0.05
sum(coh.chlaXaccndvi$accndvip.ts3<alpha)/nrow(coh.chlaXaccndvi)
sum(coh.chlaXmaxndvi$maxndvip.ts3<alpha)/nrow(coh.chlaXmaxndvi)

print(coh.chlaXaccndvi$accndviphi.ts3[coh.chlaXaccndvi$accndvip.ts3<alpha]/pi)
print(coh.chlaXmaxndvi$maxndviphi.ts3[coh.chlaXmaxndvi$maxndvip.ts3<alpha]/pi)

```

```{r add vars for modelling, echo=TRUE, cache=FALSE}

#Need to add: depth, average growing season Chlorophyll-a, TSI(chla) categories, pct ag

#agriculture -- is 500m buffer best? Other options include 100m buffer (probably too small) and hu12 watershed
pct.ag<-lagosne_select(table="buffer500m.lulc", vars=c("lagoslakeid","buffer500m_nlcd2001_pct_82","buffer500m_nlcd2006_pct_82","buffer500m_nlcd2011_pct_82"))
pct.ag<-pct.ag[pct.ag$lagoslakeid %in% analysislakes$lakeinfo$lagoslakeid,]
pct.ag.avg<-data.frame(lagoslakeid=pct.ag$lagoslakeid, pct.ag=rowMeans(pct.ag[,2:4]))

#depth
depth<-lagosne_select(table="lakes_limno", vars=c("lagoslakeid","maxdepth"))
depth<-depth[depth$lagoslakeid %in% analysislakes$lakeinfo$lagoslakeid,] #use max depth because it's more complete

#growing season Chlorophyll-a
chla<-lagosne_select(table="epi_nutr", vars=c("lagoslakeid","samplemonth","chla"))
chla<-chla[chla$lagoslakeid %in% analysislakes$lakeinfo$lagoslakeid,]
gs.chla<-chla[chla$samplemonth %in% 5:9,]
avg.chla<-aggregate(chla ~ lagoslakeid, data=gs.chla, FUN=mean, na.rm=T)

#Chlorophyll-a TSI class
#TSI(CHL) = 9.81 ln(CHL) + 30.6
tsi.chl<-data.frame(lagoslakeid=avg.chla$lagoslakeid, tsi=9.81 * log(avg.chla$chla) + 30.6)
tsi.chl$tsi.cat<-rep("lake",nrow(tsi.chl))

tsi.chl$tsi.cat[tsi.chl$tsi < 40]<-"oligotrophic"
tsi.chl$tsi.cat[tsi.chl$tsi >=40 & tsi.chl$tsi < 50]<-"mesotrophic"
tsi.chl$tsi.cat[tsi.chl$tsi >=50 & tsi.chl$tsi < 70]<-"eutrophic"
tsi.chl$tsi.cat[tsi.chl$tsi >= 70] <-"hypereutrophic"

#huc2 and huc4 watershed codes
huc_codes<-read.csv("/Users/jonathanwalter/GitHub/AquaTerrSynch/AnalysisCode/match_huc_codes.csv", colClasses = 'character')

predictors<-analysislakes$lakeinfo
predictors$tslength<-predictors$end-predictors$start+1
predictors<-left_join(predictors, depth, by="lagoslakeid")
predictors<-left_join(predictors, pct.ag.avg, by="lagoslakeid")
predictors<-left_join(predictors, avg.chla, by="lagoslakeid")
predictors<-left_join(predictors, tsi.chl, by="lagoslakeid")
predictors<-left_join(predictors, huc_codes, by="hu4_zoneid")

modvars.accndvi<-left_join(predictors, coh.chlaXaccndvi, by="lagoslakeid")
modvars.accndvi$nhd_ftype<-factor(modvars.accndvi$nhd_ftype)
modvars.accndvi$tsi.cat<-factor(modvars.accndvi$tsi.cat)
```

```{r linear modelling, echo=TRUE, cache=FALSE}

modvars.accndvi<-modvars.accndvi[!is.na(modvars.accndvi$maxdepth),]
modvars.accndvi<-modvars.accndvi[!is.na(modvars.accndvi$pct.ag),]

#short timescales
fm.coh.accndvi.st<-gls(accndvicoh.ts1 ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
dredge.coh.accndvi.st<-dredge(fm.coh.accndvi.st, beta="sd") #intercept only is best model. Disappointing.
print(dredge.coh.accndvi.st)

fm.p.accndvi.st<-gls(accndvip.ts1 ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
dredge.p.accndvi.st<-dredge(fm.p.accndvi.st, beta="sd") #intercept only is best model. Disappointing.
print(dredge.p.accndvi.st)

#long timescales
fm.coh.accndvi.lt<-gls(accndvicoh.ts2 ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
dredge.coh.accndvi.lt<-dredge(fm.coh.accndvi.lt, beta="sd") #intercept only is best model. Disappointing.
print(dredge.coh.accndvi.lt)

fm.p.accndvi.lt<-gls(accndvip.ts2 ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
dredge.p.accndvi.lt<-dredge(fm.p.accndvi.lt, beta="sd") #intercept only is best model. Disappointing.
print(dredge.p.accndvi.lt)
```