### 2019-12-10 AnalysisCode/Q2_ClimateCoherence/ms2_analyses.Rmd
q2 updates in progress

---
title: 'Q2: Does climate drive lake-landscape coherence?'
author: "Jonathan Walter, Grace Wilkinson, Rachel Fleck, Michael Pace"
date: "12/9/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(aqts)
library(wsyn)
library(rgdal)
library(rgeos)
library(LAGOSNE)
library(raster)
library(dplyr)
library(viridis)

rm(list=ls())

```

This document organizes for openness and reproducibility analyses of the influence of climate .

# Data import

In the Q1 manuscript, we output time series from coherent lake-landscape pairs. These are loaded here. We also load time series of three major climate indices (ENSO, PDO, NAO) and local temperature and precipitation.

```{r load and format data, echo=TRUE, cache=FALSE}

#load coherent lakes data
load("~/Box Sync/NSF EAGER Synchrony/Data/RData files/q2_coherent_lakes.RData")
#load climate index time series
mei.raw<-read.csv("/Users/jonathanwalter/Documents/Research/DATA/ClimateIndices/mei_monthly_1979_2019.csv")#MEI
pdo.raw<-read.csv("/Users/jonathanwalter/Documents/Research/DATA/ClimateIndices/pdo_monthly_1950_2019.csv")#PDO
nao.raw<-read.csv("/Users/jonathanwalter/Documents/Research/DATA/ClimateIndices/nao_monthly_1950_2019.csv")#NAO

mei.gs<-mei.raw[mei.raw$mm >=5 & mei.raw$mm <=9,]
pdo.gs<-pdo.raw[pdo.raw$mm >=5 & pdo.raw$mm <=9,]
nao.gs<-nao.raw[nao.raw$mm >=5 & nao.raw$mm <=9,]

mei.gs<-aggregate(mei.gs$mei,list(mei.gs$yyyy),FUN="mean")
names(mei.gs)<-c("year","mei")
pdo.gs<-aggregate(pdo.gs$PDO,list(pdo.gs$yyyy),FUN="mean")
names(pdo.gs)<-c("year","pdo")
nao.gs<-aggregate(nao.gs$NAO,list(nao.gs$yyyy),FUN="mean")
names(nao.gs)<-c("year","nao")

for(ii in 1:length(coherentlakes.st$lakedata)){
  coherentlakes.st$lakedata[[ii]]<-rbind(coherentlakes.st$lakedata[[ii]],
                                         mei.gs$mei[mei.gs$year %in% colnames(coherentlakes.st$lakedata[[ii]])],
                                         pdo.gs$pdo[pdo.gs$year %in% colnames(coherentlakes.st$lakedata[[ii]])],
                                         nao.gs$nao[nao.gs$year %in% colnames(coherentlakes.st$lakedata[[ii]])])
  rownames(coherentlakes.st$lakedata[[ii]])<-c("chla","ndvi","mei","pdo","nao")
}

for(ii in 1:length(coherentlakes.lt$lakedata)){
  coherentlakes.lt$lakedata[[ii]]<-rbind(coherentlakes.lt$lakedata[[ii]],
                                         mei.gs$mei[mei.gs$year %in% colnames(coherentlakes.lt$lakedata[[ii]])],
                                         pdo.gs$pdo[pdo.gs$year %in% colnames(coherentlakes.lt$lakedata[[ii]])],
                                         nao.gs$nao[nao.gs$year %in% colnames(coherentlakes.lt$lakedata[[ii]])])
  rownames(coherentlakes.lt$lakedata[[ii]])<-c("chla","ndvi","mei","pdo","nao")
}
#add local weather --  write a function to do this.

```

# Analyses

a) Are lakes and landscapes coherent with the same climate variables? b) Do climate variables explain a similar amount of variability in coherent lakes as the landscape does? c) does this differ by timescale?

First, do coherences with climate variables
```{r climate coherence, echo=TRUE, cache=FALSE}

```

Second, find out how much variability in lakes is explained by coherence with the landscape
```{r wlms lake v landscape, echo=TRUE, cache=FALSE}

```

Third, find out how much variability in lakes is explained by coherence with variables the landscape is coherent with
```{r wlms lake v climate, echo=TRUE, cache=FALSE}

```