### 2019-12-11 AnalysisCode/Q2_ClimateCoherence/ms2_analyses.Rmd
q2 analyses updates

---
title: 'Q2: Does climate drive lake-landscape coherence?'
author: "Jonathan Walter, Grace Wilkinson, Rachel Fleck, Michael Pace"
date: "12/9/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(aqts)
library(wsyn)
library(rgdal)
library(rgeos)
library(LAGOSNE)
library(raster)
library(dplyr)
library(viridis)

rm(list=ls())

source("/Users/jonathanwalter/GitHub/AquaTerrSynch/AnalysisCode/Q2_ClimateCoherence/bandtest_coh.R")

```

This document organizes for openness and reproducibility analyses of the influence of climate .

# Data import

In the Q1 manuscript, we output time series from coherent lake-landscape pairs. These are loaded here. We also load time series of three major climate indices (ENSO, PDO, NAO) and local temperature and precipitation.

```{r load and format data, echo=TRUE, cache=TRUE}

#load coherent lakes data
load("~/Box Sync/NSF EAGER Synchrony/Data/RData files/q2_coherent_lakes.RData")
#load climate index time series
mei.raw<-read.csv("/Users/jonathanwalter/Documents/Research/DATA/ClimateIndices/mei_monthly_1979_2019.csv")#MEI
pdo.raw<-read.csv("/Users/jonathanwalter/Documents/Research/DATA/ClimateIndices/pdo_monthly_1950_2019.csv")#PDO
nao.raw<-read.csv("/Users/jonathanwalter/Documents/Research/DATA/ClimateIndices/nao_monthly_1950_2019.csv")#NAO

mei.gs<-mei.raw[mei.raw$mm >=5 & mei.raw$mm <=9,]
pdo.gs<-pdo.raw[pdo.raw$mm >=5 & pdo.raw$mm <=9,]
nao.gs<-nao.raw[nao.raw$mm >=5 & nao.raw$mm <=9,]

mei.gs<-aggregate(mei.gs$mei,list(mei.gs$yyyy),FUN="mean")
names(mei.gs)<-c("year","mei")
pdo.gs<-aggregate(pdo.gs$PDO,list(pdo.gs$yyyy),FUN="mean")
names(pdo.gs)<-c("year","pdo")
nao.gs<-aggregate(nao.gs$NAO,list(nao.gs$yyyy),FUN="mean")
names(nao.gs)<-c("year","nao")

for(ii in 1:length(coherentlakes.st$lakedata)){
  coherentlakes.st$lakedata[[ii]]<-rbind(coherentlakes.st$lakedata[[ii]],
                                         mei.gs$mei[mei.gs$year %in% colnames(coherentlakes.st$lakedata[[ii]])],
                                         pdo.gs$pdo[pdo.gs$year %in% colnames(coherentlakes.st$lakedata[[ii]])],
                                         nao.gs$nao[nao.gs$year %in% colnames(coherentlakes.st$lakedata[[ii]])])
  rownames(coherentlakes.st$lakedata[[ii]])<-c("chla","ndvi","mei","pdo","nao")
}

for(ii in 1:length(coherentlakes.lt$lakedata)){
  coherentlakes.lt$lakedata[[ii]]<-rbind(coherentlakes.lt$lakedata[[ii]],
                                         mei.gs$mei[mei.gs$year %in% colnames(coherentlakes.lt$lakedata[[ii]])],
                                         pdo.gs$pdo[pdo.gs$year %in% colnames(coherentlakes.lt$lakedata[[ii]])],
                                         nao.gs$nao[nao.gs$year %in% colnames(coherentlakes.lt$lakedata[[ii]])])
  rownames(coherentlakes.lt$lakedata[[ii]])<-c("chla","ndvi","mei","pdo","nao")
}

#add local weather
ppt.files<-paste0("/Users/jonathanwalter/Documents/Research/DATA/PRISM_4km2_gridded/ppt/PRISM_ppt_stable_4kmM3_",
                  rep(1989:2018,each=5),
                  rep(c("05","06","07","08","09"),times=30),
                  "_bil.bil")
ppt<-stack(ppt.files)

tavg.files<-c(paste0("/Users/jonathanwalter/Documents/Research/DATA/PRISM_4km2_gridded/tmean/PRISM_tmean_stable_4kmM2_",
                  rep(1989:2016,each=5),
                  rep(c("05","06","07","08","09"),times=28),
                  "_bil.bil"),
              paste0("/Users/jonathanwalter/Documents/Research/DATA/PRISM_4km2_gridded/tmean/PRISM_tmean_stable_4kmM3_",
                  rep(2017:2018,each=5),
                  rep(c("05","06","07","08","09"),times=2),
                  "_bil.bil"))
tavg<-stack(tavg.files)

coherentlakes.st<-addPRISMts(coherentlakes.st, ppt, var="ppt")
coherentlakes.st<-addPRISMts(coherentlakes.st, tavg, var="tmean")
coherentlakes.lt<-addPRISMts(coherentlakes.lt, ppt, var="ppt")
coherentlakes.lt<-addPRISMts(coherentlakes.lt, tavg, var="tmean")
```

# Analyses

a) Are lakes and landscapes coherent with the same climate variables? b) Do climate variables explain a similar amount of variability in coherent lakes as the landscape does? c) does this differ by timescale?

First, do coherences with climate variables
```{r climate coherence, echo=TRUE, cache=FALSE}
lakeXclim.st<-data.frame(lagoslakeid=names(coherentlakes.st$lakedata),
                         mei.coh=rep(NA, length(coherentlakes.st$lakedata)),
                         mei.p=rep(NA, length(coherentlakes.st$lakedata)),
                         mei.phi=rep(NA, length(coherentlakes.st$lakedata)),
                         nao.coh=rep(NA, length(coherentlakes.st$lakedata)),
                         nao.p=rep(NA, length(coherentlakes.st$lakedata)),
                         nao.phi=rep(NA, length(coherentlakes.st$lakedata)),
                         pdo.coh=rep(NA, length(coherentlakes.st$lakedata)),
                         pdo.p=rep(NA, length(coherentlakes.st$lakedata)),
                         pdo.phi=rep(NA, length(coherentlakes.st$lakedata)),
                         ppt.coh=rep(NA, length(coherentlakes.st$lakedata)),
                         ppt.p=rep(NA, length(coherentlakes.st$lakedata)),
                         ppt.phi=rep(NA, length(coherentlakes.st$lakedata)),
                         tavg.coh=rep(NA, length(coherentlakes.st$lakedata)),
                         tavg.p=rep(NA, length(coherentlakes.st$lakedata)),
                         tavg.phi=rep(NA, length(coherentlakes.st$lakedata)))
landXclim.st<-lakeXclim.st

lakeXclim.lt<-data.frame(lagoslakeid=names(coherentlakes.lt$lakedata),
                         mei.coh=rep(NA, length(coherentlakes.lt$lakedata)),
                         mei.p=rep(NA, length(coherentlakes.lt$lakedata)),
                         mei.phi=rep(NA, length(coherentlakes.lt$lakedata)),
                         nao.coh=rep(NA, length(coherentlakes.lt$lakedata)),
                         nao.p=rep(NA, length(coherentlakes.lt$lakedata)),
                         nao.phi=rep(NA, length(coherentlakes.lt$lakedata)),
                         pdo.coh=rep(NA, length(coherentlakes.lt$lakedata)),
                         pdo.p=rep(NA, length(coherentlakes.lt$lakedata)),
                         pdo.phi=rep(NA, length(coherentlakes.lt$lakedata)),
                         ppt.coh=rep(NA, length(coherentlakes.lt$lakedata)),
                         ppt.p=rep(NA, length(coherentlakes.lt$lakedata)),
                         ppt.phi=rep(NA, length(coherentlakes.lt$lakedata)),
                         tavg.coh=rep(NA, length(coherentlakes.lt$lakedata)),
                         tavg.p=rep(NA, length(coherentlakes.lt$lakedata)),
                         tavg.phi=rep(NA, length(coherentlakes.lt$lakedata)))
landXclim.lt<-lakeXclim.lt

# lakes, short timescales
NN=length(coherentlakes.st$lakedata)
ts=c(2,4)
for(ii in 1:NN){
  vars<-rownames(coherentlakes.st$lakedata[[ii]])
  yy<-as.numeric(colnames(coherentlakes.st$lakedata[[ii]]))
  clnd<-cleandat(coherentlakes.st$lakedata[[ii]],yy,clev=5)$cdat

  cohXmei<-coh(clnd[vars=="chla",], clnd[vars=="mei"], yy, norm="powall", sigmethod="fast")
  cohXnao<-coh(clnd[vars=="chla",], clnd[vars=="nao"], yy, norm="powall", sigmethod="fast")
  cohXpdo<-coh(clnd[vars=="chla",], clnd[vars=="pdo"], yy, norm="powall", sigmethod="fast")
  cohXppt<-coh(clnd[vars=="chla",], clnd[vars=="ppt"], yy, norm="powall", sigmethod="fast")
  cohXtavg<-coh(clnd[vars=="chla",], clnd[vars=="tavg"], yy, norm="powall", sigmethod="fast")

  cohXmei<-bandtest.coh(cohXmei,ts)
  cohXnao<-bandtest.coh(cohXnao,ts)
  cohXpdo<-bandtest.coh(cohXpdo,ts)
  cohXppt<-bandtest.coh(cohXppt,ts)
  cohXtavg<-bandtest.coh(cohXtavg,ts)

  lakeXclim.st$mei.coh[ii]<-cohXmei$bandp$mn_coh
  lakeXclim.st$mei.p[ii]<-cohXmei$bandp$p_val
  lakeXclim.st$mei.phi[ii]<-cohXmei$bandp$mn_phs
  lakeXclim.st$nao.coh[ii]<-cohXnao$bandp$mn_coh
  lakeXclim.st$nao.p[ii]<-cohXnao$bandp$p_val
  lakeXclim.st$nao.phi[ii]<-cohXnao$bandp$mn_phs
  lakeXclim.st$pdo.coh[ii]<-cohXpdo$bandp$mn_coh
  lakeXclim.st$pdo.p[ii]<-cohXpdo$bandp$p_val
  lakeXclim.st$pdo.phi[ii]<-cohXpdo$bandp$mn_phs
  lakeXclim.st$ppt.coh[ii]<-cohXppt$bandp$mn_coh
  lakeXclim.st$ppt.p[ii]<-cohXppt$bandp$p_val
  lakeXclim.st$ppt.phi[ii]<-cohXppt$bandp$mn_phs
  lakeXclim.st$tavg.coh[ii]<-cohXtavg$bandp$mn_coh
  lakeXclim.st$tavg.p[ii]<-cohXtavg$bandp$p_val
  lakeXclim.st$tavg.phi[ii]<-cohXtavg$bandp$mn_phs
}

#land, short timescales
NN=length(coherentlakes.st$lakedata)
ts=c(2,4)
for(ii in 1:NN){
  vars<-rownames(coherentlakes.st$lakedata[[ii]])
  yy<-as.numeric(colnames(coherentlakes.st$lakedata[[ii]]))
  clnd<-cleandat(coherentlakes.st$lakedata[[ii]],yy,clev=5)$cdat

  cohXmei<-coh(clnd[vars=="ndvi",], clnd[vars=="mei"], yy, norm="powall", sigmethod="fast")
  cohXnao<-coh(clnd[vars=="ndvi",], clnd[vars=="nao"], yy, norm="powall", sigmethod="fast")
  cohXpdo<-coh(clnd[vars=="ndvi",], clnd[vars=="pdo"], yy, norm="powall", sigmethod="fast")
  cohXppt<-coh(clnd[vars=="ndvi",], clnd[vars=="ppt"], yy, norm="powall", sigmethod="fast")
  cohXtavg<-coh(clnd[vars=="ndvi",], clnd[vars=="tavg"], yy, norm="powall", sigmethod="fast")

  cohXmei<-bandtest.coh(cohXmei,ts)
  cohXnao<-bandtest.coh(cohXnao,ts)
  cohXpdo<-bandtest.coh(cohXpdo,ts)
  cohXppt<-bandtest.coh(cohXppt,ts)
  cohXtavg<-bandtest.coh(cohXtavg,ts)

  landXclim.st$mei.coh[ii]<-cohXmei$bandp$mn_coh
  landXclim.st$mei.p[ii]<-cohXmei$bandp$p_val
  landXclim.st$mei.phi[ii]<-cohXmei$bandp$mn_phs
  landXclim.st$nao.coh[ii]<-cohXnao$bandp$mn_coh
  landXclim.st$nao.p[ii]<-cohXnao$bandp$p_val
  landXclim.st$nao.phi[ii]<-cohXnao$bandp$mn_phs
  landXclim.st$pdo.coh[ii]<-cohXpdo$bandp$mn_coh
  landXclim.st$pdo.p[ii]<-cohXpdo$bandp$p_val
  landXclim.st$pdo.phi[ii]<-cohXpdo$bandp$mn_phs
  landXclim.st$ppt.coh[ii]<-cohXppt$bandp$mn_coh
  landXclim.st$ppt.p[ii]<-cohXppt$bandp$p_val
  landXclim.st$ppt.phi[ii]<-cohXppt$bandp$mn_phs
  landXclim.st$tavg.coh[ii]<-cohXtavg$bandp$mn_coh
  landXclim.st$tavg.p[ii]<-cohXtavg$bandp$p_val
  landXclim.st$tavg.phi[ii]<-cohXtavg$bandp$mn_phs
}

#lakes, long timescales
NN=length(coherentlakes.lt$lakedata)
ts=c(4,Inf)
for(ii in 1:NN){
  vars<-rownames(coherentlakes.lt$lakedata[[ii]])
  yy<-as.numeric(colnames(coherentlakes.lt$lakedata[[ii]]))
  clnd<-cleandat(coherentlakes.lt$lakedata[[ii]],yy,clev=5)$cdat

  cohXmei<-coh(clnd[vars=="chla",], clnd[vars=="mei"], yy, norm="powall", sigmethod="fast")
  cohXnao<-coh(clnd[vars=="chla",], clnd[vars=="nao"], yy, norm="powall", sigmethod="fast")
  cohXpdo<-coh(clnd[vars=="chla",], clnd[vars=="pdo"], yy, norm="powall", sigmethod="fast")
  cohXppt<-coh(clnd[vars=="chla",], clnd[vars=="ppt"], yy, norm="powall", sigmethod="fast")
  cohXtavg<-coh(clnd[vars=="chla",], clnd[vars=="tavg"], yy, norm="powall", sigmethod="fast")

  cohXmei<-bandtest.coh(cohXmei,ts)
  cohXnao<-bandtest.coh(cohXnao,ts)
  cohXpdo<-bandtest.coh(cohXpdo,ts)
  cohXppt<-bandtest.coh(cohXppt,ts)
  cohXtavg<-bandtest.coh(cohXtavg,ts)

  lakeXclim.lt$mei.coh[ii]<-cohXmei$bandp$mn_coh
  lakeXclim.lt$mei.p[ii]<-cohXmei$bandp$p_val
  lakeXclim.lt$mei.phi[ii]<-cohXmei$bandp$mn_phs
  lakeXclim.lt$nao.coh[ii]<-cohXnao$bandp$mn_coh
  lakeXclim.lt$nao.p[ii]<-cohXnao$bandp$p_val
  lakeXclim.lt$nao.phi[ii]<-cohXnao$bandp$mn_phs
  lakeXclim.lt$pdo.coh[ii]<-cohXpdo$bandp$mn_coh
  lakeXclim.lt$pdo.p[ii]<-cohXpdo$bandp$p_val
  lakeXclim.lt$pdo.phi[ii]<-cohXpdo$bandp$mn_phs
  lakeXclim.lt$ppt.coh[ii]<-cohXppt$bandp$mn_coh
  lakeXclim.lt$ppt.p[ii]<-cohXppt$bandp$p_val
  lakeXclim.lt$ppt.phi[ii]<-cohXppt$bandp$mn_phs
  lakeXclim.lt$tavg.coh[ii]<-cohXtavg$bandp$mn_coh
  lakeXclim.lt$tavg.p[ii]<-cohXtavg$bandp$p_val
  lakeXclim.lt$tavg.phi[ii]<-cohXtavg$bandp$mn_phs
}

#land, long timescales
NN=length(coherentlakes.lt$lakedata)
ts=c(4,Inf)
for(ii in 1:NN){
  vars<-rownames(coherentlakes.lt$lakedata[[ii]])
  yy<-as.numeric(colnames(coherentlakes.lt$lakedata[[ii]]))
  clnd<-cleandat(coherentlakes.lt$lakedata[[ii]],yy,clev=5)$cdat

  cohXmei<-coh(clnd[vars=="ndvi",], clnd[vars=="mei"], yy, norm="powall", sigmethod="fast")
  cohXnao<-coh(clnd[vars=="ndvi",], clnd[vars=="nao"], yy, norm="powall", sigmethod="fast")
  cohXpdo<-coh(clnd[vars=="ndvi",], clnd[vars=="pdo"], yy, norm="powall", sigmethod="fast")
  cohXppt<-coh(clnd[vars=="ndvi",], clnd[vars=="ppt"], yy, norm="powall", sigmethod="fast")
  cohXtavg<-coh(clnd[vars=="ndvi",], clnd[vars=="tavg"], yy, norm="powall", sigmethod="fast")

  cohXmei<-bandtest.coh(cohXmei,ts)
  cohXnao<-bandtest.coh(cohXnao,ts)
  cohXpdo<-bandtest.coh(cohXpdo,ts)
  cohXppt<-bandtest.coh(cohXppt,ts)
  cohXtavg<-bandtest.coh(cohXtavg,ts)

  landXclim.lt$mei.coh[ii]<-cohXmei$bandp$mn_coh
  landXclim.lt$mei.p[ii]<-cohXmei$bandp$p_val
  landXclim.lt$mei.phi[ii]<-cohXmei$bandp$mn_phs
  landXclim.lt$nao.coh[ii]<-cohXnao$bandp$mn_coh
  landXclim.lt$nao.p[ii]<-cohXnao$bandp$p_val
  landXclim.lt$nao.phi[ii]<-cohXnao$bandp$mn_phs
  landXclim.lt$pdo.coh[ii]<-cohXpdo$bandp$mn_coh
  landXclim.lt$pdo.p[ii]<-cohXpdo$bandp$p_val
  landXclim.lt$pdo.phi[ii]<-cohXpdo$bandp$mn_phs
  landXclim.lt$ppt.coh[ii]<-cohXppt$bandp$mn_coh
  landXclim.lt$ppt.p[ii]<-cohXppt$bandp$p_val
  landXclim.lt$ppt.phi[ii]<-cohXppt$bandp$mn_phs
  landXclim.lt$tavg.coh[ii]<-cohXtavg$bandp$mn_coh
  landXclim.lt$tavg.p[ii]<-cohXtavg$bandp$p_val
  landXclim.lt$tavg.phi[ii]<-cohXtavg$bandp$mn_phs
}

# test correlation between coherence of lakes and coherence of land
cor.test(lakeXclim.st$mei.coh,landXclim.st$mei.coh)
cor.test(lakeXclim.st$nao.coh,landXclim.st$nao.coh)
cor.test(lakeXclim.st$pdo.coh,landXclim.st$pdo.coh)
cor.test(lakeXclim.st$ppt.coh,landXclim.st$ppt.coh)
cor.test(lakeXclim.st$tavg.coh,landXclim.st$tavg.coh)
# no significant correlations at short timescales

cor.test(lakeXclim.lt$mei.coh,landXclim.lt$mei.coh)
cor.test(lakeXclim.lt$nao.coh,landXclim.lt$nao.coh)
cor.test(lakeXclim.lt$pdo.coh,landXclim.lt$pdo.coh)
cor.test(lakeXclim.lt$ppt.coh,landXclim.lt$ppt.coh)
cor.test(lakeXclim.lt$tavg.coh,landXclim.lt$tavg.coh)
# mostly significant correlations at long timescales

```
Coherence with climate variables was essentially uncorrelated between lakes and the landscape at short timescales, albeit with a small (n=9) number of lakes.
However, at long timescales coherence with climate variables tended to be significantly and positively correlated between lakes and the landscape.

Second, find out how much variability in lakes is explained by coherence with the landscape
```{r wlms lake v landscape, echo=TRUE, cache=FALSE}

## Short timescales

#2851 - Walloon Lake
yy<-as.numeric(colnames(coherentlakes.st$lakedata$`2851`))
dat<-list(cleandat(coherentlakes.st$lakedata$`2851`[1,],yy,clev=5)$cdat,
          cleandat(coherentlakes.st$lakedata$`2851`[2,],yy,clev=5)$cdat)
wlm.lakeland.2851<-wlm(dat,yy,1,2,norm="powall")
syncexpl(wlm.lakeland.2851)

```

Third, find out how much variability in lakes is explained by coherence with variables the landscape is coherent with
```{r wlms lake v climate, echo=TRUE, cache=FALSE}

```